---
title: "Lesson 4"
output: html_notebook
---

```{r}
rm(list = ls())
library(shiny)
```

Display reactive output

Time to give your Shiny app a “live” quality! This lesson will teach you how to build reactive output to displays in your Shiny app. Reactive output automatically responds when your user toggles a widget.

By the end of this lesson, you’ll know how to make a simple Shiny app with two reactive lines of text. Each line will display the values of a widget based on your user’s input.

Census visualization app

This new Shiny app will need its own, new directory. Create a folder in your working directory named census-app. This is where we’ll save the app.R file that you make in this lesson.

Two steps

You can create reactive output with a two step process.

Add an R object to your user interface.
Tell Shiny how to build the object in the server function. The object will be reactive if the code that builds it calls a widget value.
Step 1: Add an R object to the UI
Shiny provides a family of functions that turn R objects into output for your user interface. Each function creates a specific type of output.

Output function	Creates
dataTableOutput	DataTable
htmlOutput	raw HTML
imageOutput	image
plotOutput	plot
tableOutput	table
textOutput	text
uiOutput	raw HTML
verbatimTextOutput	text
You can add output to the user interface in the same way that you added HTML elements and widgets. Place the output function inside sidebarPanel or mainPanel in the ui.

For example, the ui object below uses textOutput to add a reactive line of text to the main panel of the Shiny app pictured above.
```{r}
ui <- fluidPage(
  titlePanel("censusVis"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Create demographic maps with 
               information from the 2010 US Census."),
      
      selectInput("var", 
                  label = "Choose a variable to display",
                  choices = c("Percent White", 
                              "Percent Black",
                              "Percent Hispanic", 
                              "Percent Asian"),
                  selected = "Percent White"),
      
      sliderInput("range", 
                  label = "Range of interest:",
                  min = 0, max = 100, value = c(0, 100))
    ),
    
    mainPanel(
      textOutput("selected_var")
    )
  )
)
```


Notice that textOutput takes an argument, the character string "selected_var". Each of the *Output functions require a single argument: a character string that Shiny will use as the name of your reactive element. Your users will not see this name, but you will use it later.

Step 2: Provide R code to build the object.
Placing a function in ui tells Shiny where to display your object. Next, you need to tell Shiny how to build the object.

We do this by providing the R code that builds the object in the server function.

The server function plays a special role in the Shiny process; it builds a list-like object named output that contains all of the code needed to update the R objects in your app. Each R object needs to have its own entry in the list.

You can create an entry by defining a new element for output within the server function, like below. The element name should match the name of the reactive element that you created in the ui.

In the server function below, output$selected_var matches textOutput("selected_var") in your ui.
```{r}
server <- function(input, output) {
  
  output$selected_var <- renderText({ 
    "You have selected this"
  })
  
}
```


You do not need to explicitly state in the server function to return output in its last line of code. R will automatically update output through reference class semantics.

Each entry to output should contain the output of one of Shiny’s render* functions. These functions capture an R expression and do some light pre-processing on the expression. Use the render* function that corrresponds to the type of reactive object you are making.

render function	creates
renderDataTable	DataTable
renderImage	images (saved as a link to a source file)
renderPlot	plots
renderPrint	any printed output
renderTable	data frame, matrix, other table like structures
renderText	character strings
renderUI	a Shiny tag object or HTML
Each render* function takes a single argument: an R expression surrounded by braces, {}. The expression can be one simple line of text, or it can involve many lines of code, as if it were a complicated function call.

Think of this R expression as a set of instructions that you give Shiny to store for later. Shiny will run the instructions when you first launch your app, and then Shiny will re-run the instructions every time it needs to update your object.

For this to work, your expression should return the object you have in mind (a piece of text, a plot, a data frame, etc.). You will get an error if the expression does not return an object, or if it returns the wrong type of object.

Use widget values
If you run the app with the server function above, it will display “You have selected this” in the main panel. However, the text will not be reactive. It will not change even if you manipulate the widgets of your app.

You can make the text reactive by asking Shiny to call a widget value when it builds the text. Let’s look at how to do this.

Take a look at the first line of code in the server function. Do you notice that the server function mentions two arguments, input and output? You already saw that output is a list-like object that stores instructions for building the R objects in your app.

input is a second list-like object. It stores the current values of all of the widgets in your app. These values will be saved under the names that you gave the widgets in your ui.

So for example, our app has two widgets, one named "var" and one named "range" (you gave the widgets these names in Lesson 3). The values of "var" and "range" will be saved in input as input$var and input$range. Since the slider widget has two values (a min and a max), input$range will contain a vector of length two.

Shiny will automatically make an object reactive if the object uses an input value. For example, the server function below creates a reactive line of text by calling the value of the select box widget to build the text.
```{r}
server <- function(input, output) {
  
  output$selected_var <- renderText({ 
    paste("You have selected", input$var)
  })
  
}
```


Shiny tracks which outputs depend on which widgets. When a user changes a widget, Shiny will rebuild all of the outputs that depend on the widget, using the new value of the widget as it goes. As a result, the rebuilt objects will be completely up-to-date.

This is how you create reactivity with Shiny, by connecting the values of input to the objects in output. Shiny takes care of all of the other details.

Launch your app and see the reactive output

When you are ready, update your server and ui functions to match those above. Then launch your Shiny app by running runApp("census-app", display.mode = "showcase") at the command line. Your app should look like the app below, and your statement should update instantly as you change the select box widget.
```{r}
runApp("census-app", display.mode = "showcase")
#Do not know how to create new directory so this is not working for me
```


Watch the server portion of the script. When Shiny rebuilds an output, it highlights the code it is running. This temporary highlighting can help you see how Shiny generates reactive output.

Census visualization - showcase

Your turn

Add a second line of reactive text to the main panel of your Shiny app. This line should display “You have chosen a range that goes from something to something”, and each something should show the current minimum (min) or maximum (max) value of the slider widget.

Don’t forget to update both your ui object and your server function.

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("censusVis"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Create demographic maps with 
               information from the 2010 US Census."),
      
      selectInput("var", 
                  label = "Choose a variable to display",
                  choices = c("Percent White", 
                              "Percent Black",
                              "Percent Hispanic", 
                              "Percent Asian"),
                  selected = "Percent White"),
      
      sliderInput("range", 
                  label = "Range of interest:",
                  min = 0, max = 100, value = c(0, 100))
    ),
    
    mainPanel(
      textOutput("selected_var"),
      textOutput("min_max")
    )
  )
)

server <- function(input, output) {
  
  output$selected_var <- renderText({ 
    paste("You have selected", input$var)
  })
  
  output$min_max <- renderText({ 
    paste("You have chosen a range that goes from",
          input$range[1], "to", input$range[2])
  })
  
}

shinyApp(ui, server)
```

Recap

In this lesson, you created your first reactive Shiny app. Along the way, you learned to

use an *Output function in the ui to place reactive objects in your Shiny app,
use a render* function in the server to tell Shiny how to build your objects,
surround R expressions by curly braces, {}, in each render* function,
save your render* expressions in the output list, with one entry for each reactive object in your app, and
create reactivity by including an input value in a render* expression.
If you follow these rules, Shiny will automatically make your objects reactive.

