---
title: "Shiny Lesson 1"
output: html_notebook
---

The purpose of this R Notebook is to go through the Shiny app tutorial Lesson 1 found on this url https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/.

```{r}
rm(list = ls())
install.packages("shiny")
library(shiny)
library(httpuv)
install.packages(c("Rcpp", "httpuv", "shiny"))
install.packages("httpuv")
```

The Hello Shiny example plots a histogram of R’s faithful dataset with a configurable number of bins. Users can change the number of bins with a slider bar, and the app will immediately respond to their input. You’ll use Hello Shiny to explore the structure of a Shiny app and to create your first app.

To run Hello Shiny, type:

```{r}
library(shiny)
runExample("01_hello")
```


Structure of a Shiny App

Shiny apps are contained in a single script called app.R. The script app.R lives in a directory (for example, newdir/) and the app can be run with runApp("newdir").

app.R has three components:

a user interface object

a server function

a call to the shinyApp function

The user interface (ui) object controls the layout and appearance of your app. The server function contains the instructions that your computer needs to build your app. Finally the shinyApp function creates Shiny app objects from an explicit UI/server pair.

Note: Prior to version 0.10.2, Shiny did not support single-file apps and the ui object and server function needed to be contained in separate scripts called ui.R and server.R, respectively. This functionality is still supported in Shiny, however the tutorial and much of the supporting documentation focus on single-file apps.

One nice feature about single-file apps is that you can copy and paste the entire app into the R console, which makes it easy to quickly share code for others to experiment with. For example, if you copy and paste the code above into the R command line, it will start a Shiny app.

ui
Here is the ui object for the Hello Shiny example.
```{r}
library(shiny)

# Define UI for app that draws a histogram ----
ui <- fluidPage(

  # App title ----
  titlePanel("Hello Shiny!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)
```


Here is the server function for the Hello Shiny example.
```{r}

# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")

    })

}
```

At one level, the Hello Shiny server function is very simple. The script does some calculations and then plots a histogram with the requested number of bins.

However, you’ll also notice that most of the script is wrapped in a call to renderPlot. The comment above the function explains a bit about this, but if you find it confusing, don’t worry. We’ll cover this concept in much more detail soon.

Play with the Hello Shiny app and review the source code. Try to develop a feel for how the app works. But before you do so, note that in your app.R file you will need to start with loading the Shiny package and end with a call to shinyApp:
```{r}
library(shiny)

# See above for the definitions of ui and server
ui <- fluidPage(

  # App title ----
  titlePanel("Testing Shiny!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)

server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Testing xlab",
         main = "Histogram of waiting times")

    })

}

shinyApp(ui = ui, server = server)
```


Your R session will be busy while the Hello Shiny app is active, so you will not be able to run any R commands. R is monitoring the app and executing the app’s reactions. To get your R session back, hit escape or click the stop sign icon (found in the upper right corner of the RStudio console panel).

Running an App

Every Shiny app has the same structure: an app.R file that contains ui and server. You can create a Shiny app by making a new directory and saving an app.R file inside it. It is recommended that each app will live in its own unique directory.

You can run a Shiny app by giving the name of its directory to the function runApp. For example if your Shiny app is in a directory called my_app, run it with the following code:
```{r}
library(shiny)
runApp("my_app")
```


Note: runApp is similar to read.csv, read.table, and many other functions in R. The first argument of runApp is the filepath from your working directory to the app’s directory. The code above assumes that the app directory is in your working directory. In this case, the filepath is just the name of the directory.

(In case you are wondering, the Hello Shiny app’s files are saved in a special system directory called "01_hello". This directory is designed to work with the runExample ("01_hello") call.)

Your Turn

Create a new directory named App-1 in your working directory. Then copy and paste the app.R script above into your directory (the scripts from Hello Shiny). 

```{r}

# See above for the definitions of ui and server
ui <- fluidPage(

  # App title ----
  titlePanel("Hello World!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 5,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)

server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "orange",
         xlab = "Testing xlab",
         main = "Histogram of waiting times")

    })

}

shinyApp(ui = ui, server = server)
```


Launch your app by running runApp("App-1"). Then click escape and make some changes to your app:

Change the title from “Hello Shiny!” to “Hello World!”.

Set the minimum value of the slider bar to 5.

Change the histogram border color from "white" to "orange".

When you are ready, launch your app again. Your new app should match the image below. If it doesn’t, or if you want to check your code, press the model answers button to reveal how we did these tasks.

