---
title: "Shiny App Species"
output: html_notebook
---



This R Notebook will contain all the shiny apps broken down by species Hoping to make tabs with visual results based on species.


First, let's clear the environment and load the shiny package.
```{r}
rm(list = ls())
library(shiny)
library(dplyr)
library(readr)
library(ggplot2)
library(tidyverse)
library(leaflet)
library(maptools)
library(rgdal)
library(stringr)
```
Now let's load the data frame we need. This will tell us the total number of species per forest and how many of each species were detected
```{r}
dataFinal <- read_csv("~/Documents/Kate_ThesisCode/dataFinal.csv")
```

Subset data to remove "Nothing Here" choice
```{r}
dump<-which(dataFinal$choice == "NOTHINGHERE")
newData<-dataFinal[-dump,]
```

```{r}

#Load shape file 
Forests<-readOGR("Data/Study Forest Locations.shp", layer = "Study Forest Locations")

#Project shape file
Forests_proj<-spTransform(Forests, CRS("+proj=longlat +datum=WGS84"))

#reconfigure data frame
dat_sum<-dataFinal %>% group_by(ForestName, choice) %>% summarise(
  number_det = n()
)
```




Shiny app showing which forest types detected which mammals
```{r}
ui <- fluidPage(

  # App title ----
  titlePanel("Mammal Distribution by Forset Type"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      selectInput("species", h3("Choose your Species"),
                  choices = c("White-tailed Deer", "Chipmunk", "Coyote", "Fisher", "Raccoon", "Red Squirrel", "Other Small Mammal", "Gray Squirrel", "Black Bear", "Red Fox", "Porcupine", "Bobcat", "Opossum", "Weasel", "Striped Skunk", "Flying Squirrel", "Snowshoe Hare", "River Otter", "Mink"), selected = "White-tailed Deer"),
        

 img(src = "nun_SLU.jpg", height = 100, width = 200)
    ),
     
    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "foresthist"),
      imageOutput(outputId = "image")
    

    )
  )
)

server <- function(input, output) {

  # Histogram of which forest types each species is found at ----

  output$foresthist <- renderPlot({

  data<-switch(input$species, 
                      "White-tailed Deer" = newData %>% filter(choice=="DEERWHITETAILED"),
                      "Chipmunk" = newData %>% filter(choice=="CHIPMUNK"),
                  "Coyote" = newData %>% filter(choice == "COYOTE"),
                  "Fisher" = newData %>% filter(choice == "FISHER"),
                  "Raccoon" = newData %>% filter(choice == "RACCOON"),
                  "Red Squirrel" = newData %>% filter(choice == "SQUIRRELRED"),
               "Gray Squirrel" = newData %>% filter(choice == "SQUIRRELGRAY"),
               "Black Bear" = newData %>% filter(choice == "BLACKBEAR"),
               "Red Fox" = newData %>% filter(choice == "FOXRED"),
               "Porcupine" = newData %>% filter(choice == "PORCUPINE"),
               "Bobcat" = newData %>% filter(choice == "BOBCAT"),
               "Weasel" = newData %>% filter(choice == "WEASEL"),
               "Striped Skunk" = newData %>% filter(choice == "SKUNKSTRIPED"),
               "Flying Squirrel" = newData %>% filter(choice == "SQUIRRELFLYING"),
               "Snowshoe Hare" = newData %>% filter(choice == "SNOWSHOEHARE"),
               "River Otter" = newData %>% filter(choice == "RIVEROTTER"),
               "Mink" = newData %>% filter(choice == "MINK"),
               "Other Small Mammal" = newData %>% filter(choice == "OTHERSMALLMAMMAL"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"),)
  
    ggplot(data, aes(ForestName)) + 
     geom_histogram(stat = "count", position = "dodge") + 
     theme_bw() + 
    xlab("Forest Type") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    
 
      })
  
  output$image <- renderImage({    
    if(input$species == "White-tailed Deer"){            
      list(src = "Deer.jpg", height = 240, width = 300)
    }                                        
    else if(input$species == "Chipmunk"){
      list(src = "Chipmunk.jpg", height = 240, width = 300)
    }
    else if(input$species == "Coyote"){
      list(src = "Coyote.jpg", height = 240, width = 300)
    }
    else if(input$species == "Fisher"){
      list(src = "Fisher.jpg", height = 240, width = 300)
    }
    else if(input$species == "Raccoon"){
      list(src = "Raccoon.jpg", height = 240, width = 300)
    }
    else if(input$species == "Red Squirrel"){
      list(src = "RedSquirrel.jpg", height = 240, width = 300)
    }
    else if(input$species == "Other Small Mammal"){
      list(src = "Mole.jpg", height = 240, width = 300)
    }
    else if(input$species == "Gray Squirrel"){
      list(src = "GraySquirrel.jpg", height = 240, width = 300)
    }
    else if(input$species == "Black Bear"){
      list(src = "blackbear.jpg", height = 240, width = 300)
    }
    else if(input$species == "Red Fox"){
      list(src = "RedFox.jpg", height = 240, width = 300)
    }
    else if(input$species == "Porcupine"){
      list(src = "Porcupine.jpg", height = 240, width = 300)
    }
    else if(input$species == "Bobcat"){
      list(src = "Bobcat.jpg", height = 240, width = 300)
    }
    else if(input$species == "Opossum"){
      list(src = "Opossum.jpg", height = 240, width = 300)
    }
    else if(input$species == "Weasel"){
      list(src = "Weasel.jpg", height = 240, width = 300)
    }
    else if(input$species == "Striped Skunk"){
      list(src = "StripedSkunk.jpg", height = 240, width = 300)
    }
    else if(input$species == "Flying Squirrel"){
      list(src = "FlyingSquirrel.jpg", height = 240, width = 300)
    }
    else if(input$species == "Snowshoe Hare"){
      list(src = "SnowshoeHare.jpg", height = 240, width = 300)
    }
    else if(input$species == "River Otter"){
      list(src = "RiverOtter.jpg", height = 240, width = 300)
    }
    else if(input$species == "Mink"){
      list(src = "Mink.jpg", height = 240, width = 300)
    }
  })

    }



shinyApp(ui = ui, server = server)
```

Make a Shiny app that shows a map of detections by species (Don't know which shape file to pull in!)
```{r}
ui <- fluidPage(

  # App title ----
  titlePanel("Mammal Distribution by Forset Type"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      selectInput("species", h3("Choose your Species"),
                  choices = c("White-tailed Deer", "Chipmunk", "Coyote", "Fisher", "Raccoon", "Red Squirrel", "Other Small Mammal", "Gray Squirrel", "Black Bear", "Red Fox", "Porcupine", "Bobcat", "Opossum", "Weasel", "Striped Skunk", "Flying Squirrel", "Snowshoe Hare", "River Otter", "Mink"), selected = "White-tailed Deer"),
        
),
    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
     leafletOutput("speciesmap")

    )
  )
)


#define server logic ----
server<-function(input, output) {

  output$speciesmap<-renderLeaflet({
    
    data<-switch(input$species, 
                      "White-tailed Deer" = dat_sum %>% filter(choice == "DEERWHITETAILED"),
                "Chipmunk" = dat_sum %>% filter(choice == "CHIPMUNK"),
               "Coyote" = dat_sum %>% filter(choice == "COYOTE"),
                "Fisher" = dat_sum %>% filter(choice == "FISHER"),
                "Raccoon" = dat_sum %>% filter(choice == "RACCOON"),
                "Red Squirrel" = dat_sum %>% filter(choice == "SQUIRRELRED"),
              "Gray Squirrel" = dat_sum %>% filter(choice == "SQUIRRELGRAY"),
              "Black Bear" = dat_sum %>% filter(choice == "BLACKBEAR"),
              "Red Fox" = dat_sum %>% filter(choice == "FOXRED"),
              "Porcupine" = dat_sum %>% filter(choice == "PORCUPINE"),
              "Bobcat" = dat_sum %>% filter(choice == "BOBCAT"),
              "Weasel" = dat_sum %>% filter(choice == "WEASEL"),
              "Striped Skunk" = dat_sum %>% filter(choice == "SKUNKSTRIPED"),
              "Flying Squirrel" = dat_sum %>% filter(choice == "SQUIRRELFLYING"),
              "Snowshoe Hare" = dat_sum %>% filter(choice == "SNOWSHOEHARE"),
              "River Otter" = dat_sum %>% filter(choice == "RIVEROTTER"),
              "Mink" = dat_sum %>% filter(choice == "MINK"),
              "Other Small Mammal" = dat_sum %>% filter(choice == "OTHERSMALLMAMMAL"),
              "Opossum" = dat_sum %>% filter(choice == "OPOSSUM"),)

    #Join data to shape file
Forests_proj@data <- left_join(Forests_proj@data, data, by = c("ForestCode"= "ForestName"))

pal <- colorNumeric("Blues", domain= data$number_det) 

labels<-sprintf( "%s, Number of Detections %s", 
                 Forests$ForestCode,Forests_proj$number_det) %>% lapply(htmltools::HTML)
  
leaflet() %>% addTiles() %>% 
  setView(lng = -75.169395, lat = 44.595466, zoom = 8) %>% 
  addPolygons(
     data = Forests_proj, 
     fillColor = ~pal(data$number_det),
     fillOpacity = 0.5,
    weight = 1, 
    col = 'red',
    highlight = highlightOptions(#highlight lets you mouse over a county and have it change color
      weight = 5,
      color = "orange", 
      bringToFront = T),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")
    ) 

})

}

# Run the app ----
shinyApp(ui = ui, server = server)
```