---
title: "Shiny App Forest"
output: html_notebook
---


This R Notebook will contain all the shiny apps broken down by forest. Hoping to make tabs with visual results based on forest name.

NOTE: Need to run the whole thing from .R file not .Rmd file because the images won't load unless run through .R file.


First, let's clear the environment and load the shiny package.
```{r}
rm(list = ls())
library(shiny)
library(dplyr)
library(readr)
library(ggplot2)
library(tidyverse)
library(htmltools)
```

Now let's load the data frame we need. This will tell us the total number of species per forest and how many of each species were detected
```{r}
dataFinal <- read_csv("~/Documents/Kate_ThesisCode/dataFinal.csv")
dataFinal$Forest <- with(dataFinal, ifelse(
  ForestName == "SH", 'South Hammond', ifelse(
  ForestName == "BC", 'Beaver Creek', ifelse(
 ForestName == "DON", 'Donnerville', ifelse(
  ForestName == "WHIP", 'Whippoorwill Corners', ifelse(
    ForestName == "WF", 'Whiskey Flats', ifelse(
      ForestName == "DEG", 'Degrasse', 'whoops')))))))

#Make new column with formal species names for labeling in graph
dataFinal$Species <- with(dataFinal, ifelse(
  choice == "DEERWHITETAILED", 'White-tailed Deer', ifelse(
  choice == "HUMAN", 'Human', ifelse(
 choice == "CHIPMUNK", 'Chipmunk', ifelse(
  choice == "COYOTE", 'Coyote', ifelse(
    choice == "SOMETHINGHERE", 'Something Here', ifelse(
      choice == "FISHER", 'Fisher', ifelse(
      choice == "COTTONTAILRABBIT", 'Cottontail Rabbit', ifelse(
        choice == "NOTHINGHERE", 'Nothing Here', ifelse(
    choice == "RACCOON", 'Raccoon', ifelse(
       choice == "GROUSERUFFED", 'Ruffed Grouse', ifelse(
     choice == "SQUIRRELRED", 'Red Squirrel', ifelse(
choice == "BLUEJAY", 'Blue Jay', ifelse(
   choice == "OTHERSMALLMAMMAL", 'Other Small Mammal', ifelse(
    choice == "OTHERBIRD", 'Other Bird', ifelse(
       choice == "OPOSSUM", 'Opossum ', ifelse(
        choice == "SQUIRRELGRAY", 'Gray Squirrel', ifelse(
       choice == "BLACKBEAR", 'Black Bear', ifelse(
       choice == "TURKEY", 'Turkey', ifelse(
   choice == "DOMESTICCAT", 'Domestic Cat', ifelse(
       choice == "DOMESTICDOG", 'Domestic Dog', ifelse(
      choice == "FOXRED", 'Red Fox', ifelse(
     choice == "PORCUPINE", 'Porcupine', ifelse(
    choice == "BOBCAT", 'Bobcat', ifelse(
      choice == "WEASEL", 'Weasel', ifelse(
     choice == "SKUNKSTRIPED", 'Striped Skunk', ifelse(
       choice == "SQUIRRELFLYING", 'Flying Squirrel', ifelse(
         choice == "SNOWSHOEHARE", 'Snowshoe Hare', ifelse(
           choice == "CROWORRAVEN", 'Crow or Raven', ifelse(
             choice == "OWL", 'Owl', ifelse(
               choice == "GOOSEWILD", 'Wild Goose', ifelse(
                 choice == "RIVEROTTER", 'River Otter', ifelse(
                   choice == "MINK", 'Mink', 'whoops')))))))))))))))))))))))))))))))))

write.csv(dataFinal, file = "dataFinal.csv")
```

Subset data to remove "Nothing Here" choice
```{r}
dump<-which(dataFinal$choice == "NOTHINGHERE")
newData<-dataFinal[-dump,]

write.csv(newData, file = "newData.csv")
```


Data for trophic tab
```{r}
mammals<- read_csv("mammals.csv")
mammals<- mammals[,2:16]
```



Data for covariates tab
```{r}
Covariates <- read_csv("/Users/kateandy/Documents/Shiny/Covariates.csv")

Cov_sum<-Covariates %>% group_by(Forest) %>% summarise(
  Water = mean(P_Water50), Mixed = mean(P_Mixed50), Evergreen = mean(P_Evergreen50), Deciduous = mean(P_Deciduous50), Development = mean(P_Development50), Agriculture = mean(P_Agriculture50), Barren = mean(P_Barren50), Shrub = mean(P_Shrub50), Wetland = mean(P_Wetland50), Herbaceous = mean(P_Herb50)
)

require(reshape2)
dat_molten <- melt(Cov_sum,"Forest")
require(ggplot2)

copy<-dat_molten
copy[is.na(copy)] = 0
all_forests<- copy %>% group_by(variable) %>% summarise(
value = mean(value)
)

write.csv(dat_molten, file = "dat_molten.csv")
write.csv(all_forests, file = "all_forests.csv")
```

For diversity shiny app
```{r}
#Load the data
Diversity_total <- read_csv("~/Documents/Kate_ThesisCode/Diversity_total.csv")

#Modify the data to match that from my diversity data in thesis code
Diversity_total$BC<-replace(Diversity_total$BC, is.na(Diversity_total$BC), 0)

#Get rid of season
Diversity_total<- Diversity_total[,-2]


divNew<- Diversity_total %>% pivot_longer(-Diversity_Indices, names_to = "Forest", values_to = "index")

divFinal<- aggregate(divNew[, -c(1:2)], by = list(divNew$Diversity_Indices, divNew$Forest), FUN = mean)

#rename from Group.1 and Group.2 
names(divFinal) <- c("Diversity_Indices", "Forest", "Index")

#Add formal name column for graph legend
divFinal$Diversity_Index <- with(divFinal, ifelse(
  Diversity_Indices == "SR", 'Species Richness', ifelse(
  Diversity_Indices == "inv_simp", 'Inverse Simpson Index', ifelse(
 Diversity_Indices == "div_shan", 'Shannon Index', 'whoops'))))

write.csv(divFinal, file = "divFinal.csv")

```

```{r}
#Load shape file 
Forests<-readOGR("Data/Shapefiles/Study Forest Locations.shp", layer = "Study Forest Locations")

#Project shape file
Forests_proj<-spTransform(Forests, CRS("+proj=longlat +datum=WGS84"))

#Add column with unabbreviated forest name
Forests@data$Forest <- with(Forests@data, ifelse(
  ForestCode == "SH", 'South Hammond', ifelse(
  ForestCode == "BC", 'Beaver Creek', ifelse(
 ForestCode == "DON", 'Donnerville', ifelse(
  ForestCode == "WHIP", 'Whippoorwill Corners', ifelse(
    ForestCode == "WF", 'Whiskey Flats', ifelse(
      ForestCode == "DEG", 'Degrasse', 'whoops')))))))

#reconfigure data frame
dat_sum<-dataFinal %>% group_by(Forest, choice) %>% summarise(
  number_det = n()
)
write.csv(dat_sum, file = "dat_sum.csv")

labels<-sprintf( "%s", 
                 Forests$Forest) %>% lapply(htmltools::HTML)
```


Combined shiny app into tabs
```{r}
ui <- fluidPage(

  # App title ----
  titlePanel("North Country Wild Zooniverse Project"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      selectInput("forest", h3("Choose your Forest"),
                  choices = c("All Forests","South Hammond", "Donnerville", "Beaver Creek", "Whippoorwill Corners", "Whiskey Flats", "Degrasse"), selected = "All Forests"),
        
     conditionalPanel( condition = "NOTHINGHERE",
                    checkboxInput("remove", "Remove Nothing Here")),
     h6("Powered by:"),
      img(src = "nun_SLU.jpg", height = 60, width = 200),

  ),
  
    

    # Main panel for displaying outputs ----
    mainPanel(  tabsetPanel(
      tabPanel("Map of Study Sites", leafletOutput(outputId = "speciesmap")),
        tabPanel("Species per Forest",
                 plotOutput(outputId = "foresthist")),
        tabPanel("Species Trophic Levels", plotOutput(outputId = "trophic")),
        tabPanel("Forest Composition", plotOutput(outputId = "covariates")),
        tabPanel("Forest Diversity", plotOutput(outputId = "diversity"))
      )
  )
)
)


server <- function(input, output) {

  # Histogram of the number of species per forest ----

 output$speciesmap <- renderLeaflet({ 
 
     
  leaflet() %>% addTiles() %>% 
  setView(lng = -75.169395, lat = 44.595466, zoom = 9) %>%  
  addPolygons(
     data = Forests_proj, 
    weight = 2, 
    col = '#39541e',
    highlight = highlightOptions(#highlight lets you mouse over a county and have it change color
      weight = 5,
      color = "#C6ABE1", 
      bringToFront = T),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")
    ) 
 
})
  output$foresthist <- renderPlot({

   
   if(input$remove){
  newdata<-switch(input$forest, 
                  "All Forests" = newData,
                      "South Hammond" = newData %>% filter(ForestName=="SH"),
                      "Beaver Creek" = newData %>% filter(ForestName=="BC"),
                  "Donnerville" = newData %>% filter(ForestName == "DON"),
                  "Whippoorwill Corners" = newData %>% filter(ForestName == "WHIP"),
                  "Whiskey Flats" = newData %>% filter(ForestName == "WF"),
                  "Degrasse" = newData %>% filter(ForestName == "DEG"))
  
    ggplot(newdata, aes(Species)) + 
     geom_histogram(stat = "count", position = "dodge", fill = '#165970', colour = '#543b1f') + 
      labs(title = "Species per Forest", x="Species", y="Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
}    
    else {
data<- switch(input$forest, 
              "All Forests" = dataFinal,
                      "South Hammond" = dataFinal %>% filter(ForestName=="SH"),
                      "Beaver Creek" = dataFinal %>% filter(ForestName=="BC"),
                  "Donnerville" = dataFinal %>% filter(ForestName == "DON"),
                  "Whippoorwill Corners" = dataFinal %>% filter(ForestName == "WHIP"),
                  "Whiskey Flats" = dataFinal %>% filter(ForestName == "WF"),
                  "Degrasse" = dataFinal %>% filter(ForestName == "DEG"))
          
      ggplot(data, aes(Species)) + 
     geom_histogram(stat = "count", position = "dodge", fill = '#165970', colour = '#543b1f') + 
        labs(title = "Species per Forest", x="Species", y="Number of Detections")+
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    }
      })
  output$trophic <- renderPlot({

  data<-switch(input$forest, 
               "All Forests" = mammals,
                      "Whiskey Flats" = mammals %>% filter(ForestName=="WF"),
                      "South Hammond" = mammals %>% filter(ForestName=="SH"),
                  "Donnerville" = mammals %>% filter(ForestName == "DON"),
                  "Beaver Creek" = mammals %>% filter(ForestName == "BC"),
                  "Degrasse" = mammals %>% filter(ForestName == "DEG"),
                  "Whippoorwill Corners" = mammals %>% filter(ForestName == "WHIP"),)
  
    ggplot(data, aes(Trophic)) + 
     geom_histogram(stat = "count", position = "dodge", fill = '#165970', colour = '#543b1f') + 
      labs(title = "Trophic Levels per Forest", x="Trophic Level", y="Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    
 
      })
  output$covariates <- renderPlot({

  data<-switch(input$forest, 
               "All Forests" = all_forests,
                      "South Hammond" = dat_molten %>% filter(Forest=="SH"),
                      "Beaver Creek" = dat_molten %>% filter(Forest=="BC"),
                  "Donnerville" = dat_molten %>% filter(Forest=="DON"),
                  "Degrasse" = dat_molten %>% filter(Forest=="DEG"),
                  "Whippoorwill Corners" = dat_molten %>% filter(Forest=="WHIP"),
                  "Whiskey Flats" = dat_molten %>% filter(Forest=="WF"),)
  
  
   ggplot(data, aes(x="", y=value, fill= variable)) + 
         geom_bar(stat="identity") + 
  coord_polar("y", start = 0) + 
     geom_text(aes(label = paste0(round(value), "%")), position = position_stack(vjust = 0.5))+
      theme_void() +
      theme(axis.line = element_blank(),
          plot.title = element_text(hjust=0.5)) + 
    labs(x = NULL,
         y = NULL,
         title = "Forest Composition")+
     scale_fill_manual(values = c("Water" = "#165970",
                                  "Mixed" = "#543b1f",
                                  "Evergreen" = "#C6ABE1",
                                  "Deciduous" = "#39541e",
                                  "Development" = "#ABC4E0",
                                  "Agriculture" = "#E9A68F",
                                   "Barren" = "#b9da97",
                                   "Shrub" = "#d4b18a",
                                  "Wetland" = "#69c3e1",
                                  "Herbaceous" = "Gray"))
    
 
      })
    output$diversity <- renderPlot({

  data<-switch(input$forest, 
                "All Forests" = divFinal,
                      "South Hammond" = divFinal %>% filter(Forest == "SH"),
                      "Beaver Creek" = divFinal %>% filter(Forest == "BC"),
                  "Donnerville" = divFinal %>% filter(Forest == "DON"),
                  "Degrasse" = divFinal %>% filter(Forest == "DEG"),
                  "Whippoorwill Corners" = divFinal %>% filter(Forest == "WHIP"),
                  "Whiskey Flats" = divFinal %>% filter(Forest == "WF"),)
  
 ggplot(data, aes(x= Forest, y = Index, fill = Diversity_Index)) + 
     geom_bar(stat = "identity",position= position_dodge(), width = 0.7) +
     labs(title = "Diversity Indices per Forest", x= "Forest", y= "Diversity Index") + theme (plot.title =element_text(hjust = 0.5)) + 
   scale_fill_manual(values = c("Shannon Index" = "#165970",
                                  "Inverse Simpson Index" = "#543b1f",
                                  "Species Richness" = "#C6ABE1"))
 
      })


 

    }



shinyApp(ui = ui, server = server)
```

```{r}
ui <- fluidPage(

  # App title ----
  titlePanel("North Country Wild Zooniverse Project"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      selectInput("forest", h3("Choose your Forest"),
                  choices = c("South Hammond", "Donnerville", "Beaver Creek", "Whippoorwill Corners", "Whiskey Flats", "Degrasse"), selected = "South Hammond"),
      
     checkboxGroupInput("checkGroup", 
                              h3("Choose your Species"), 
                              choices = list("White-tailed Deer",
                                             "Chipmunk",
                                             "Coyote",
                                             "Fisher",
                                             "Raccoon",
                                             "Red Squirrel",
                                             "Other Small Mammal",
                                             "Gray Squirrel",
                                             "Black Bear",
                                             "Red Fox",
                                             "Porcupine",
                                             "Bobcat",
                                             "Opossum",
                                             "Weasel",
                                             "Striped Skunk",
                                             "Flying Squirrel",
                                             "Snowshoe Hare",
                                             "River Otter",
                                             "Mink"),
                              selected = "White-tailed Deer")),
  

    # Main panel for displaying outputs ----
    mainPanel(
        plotOutput(outputId = "foresthist"),
      img(src = "/Users/kateandy/Documents/Shiny/www/NatureUpNorth.png", height = 100, width = 300),
)
)
)

server <- function(input, output) {
  
  

  output$foresthist <- renderPlot({
     study <- switch(input$forest,
                   "South Hammond" = newData %>% filter(ForestName=="SH"),
                      "Beaver Creek" = newData %>% filter(ForestName=="BC"),
                  "Donnerville" = newData %>% filter(ForestName == "DON"),
                  "Whippoorwill Corners" = newData %>% filter(ForestName == "WHIP"),
                  "Whiskey Flats" = newData %>% filter(ForestName == "WF"),
                  "Degrasse" = newData %>% filter(ForestName == "DEG"))
  
     data<- switch(input$species, 
                      "White-tailed Deer" = study %>% filter(choice=="DEERWHITETAILED"),
                      "Chipmunk" = study %>% filter(choice=="CHIPMUNK"),
                  "Coyote" = study %>% filter(choice == "COYOTE"),
                  "Fisher" = study %>% filter(choice == "FISHER"),
                  "Raccoon" = study %>% filter(choice == "RACCOON"),
                  "Red Squirrel" = study %>% filter(choice == "SQUIRRELRED"),
               "Gray Squirrel" = study %>% filter(choice == "SQUIRRELGRAY"),
               "Black Bear" = study %>% filter(choice == "BLACKBEAR"),
               "Red Fox" = study %>% filter(choice == "FOXRED"),
               "Porcupine" = study %>% filter(choice == "PORCUPINE"),
               "Bobcat" = study %>% filter(choice == "BOBCAT"),
               "Weasel" = study %>% filter(choice == "WEASEL"),
               "Striped Skunk" = study %>% filter(choice == "SKUNKSTRIPED"),
               "Flying Squirrel" = study %>% filter(choice == "SQUIRRELFLYING"),
               "Snowshoe Hare" = study %>% filter(choice == "SNOWSHOEHARE"),
               "River Otter" = study %>% filter(choice == "RIVEROTTER"),
               "Mink" = study %>% filter(choice == "MINK"),
               "Other Small Mammal" = study %>% filter(choice == "OTHERSMALLMAMMAL"),
               "Opossum" = study %>% filter(choice == "OPOSSUM"),)
 

          
      ggplot(data, aes("")) +
     geom_histogram(stat = "count", position = "dodge") +
     theme_bw() +
    xlab("Species") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    })

  }
shinyApp(ui = ui, server = server)
```

```{r}
ui <- fluidPage(

  # App title ----
  titlePanel("North Country Wild Zooniverse Project"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      selectInput("forest", h3("Choose your Forest"),
                  choices = c("SH", "Donnerville", "BC", "WHIP", "WF", "DEG"), selected = "SH"),
      
     checkboxGroupInput("species", 
                              h3("Choose your Species"), 
                              choices = list("White-tailed Deer",
                                             "Chipmunk",
                                             "Coyote",
                                             "Fisher",
                                             "Raccoon",
                                             "Red Squirrel",
                                             "Other Small Mammal",
                                             "Gray Squirrel",
                                             "Black Bear",
                                             "Red Fox",
                                             "Porcupine",
                                             "Bobcat",
                                             "Opossum",
                                             "Weasel",
                                             "Striped Skunk",
                                             "Flying Squirrel",
                                             "Snowshoe Hare",
                                             "River Otter",
                                             "Mink"),
                              selected = "White-tailed Deer")),
  

    # Main panel for displaying outputs ----
    mainPanel(
        plotOutput(outputId = "foresthist"),
      img(src = "NatureUpNorth.png", height = 100, width = 300),
)
)
)

server <- function(input, output) {
  
  
 

  output$foresthist <- renderPlot({
     choices<-input$species
  data<-newData %>% filter(choice %in% choices)
    study<- reactive(
    switch(input$forest,
                   "South Hammond" = data %>% filter(ForestName=="SH"),
                      "Beaver Creek" = data %>% filter(ForestName=="BC"),
                  "Donnerville" = data %>% filter(ForestName == "DON"),
                  "Whippoorwill Corners" = data %>% filter(ForestName == "WHIP"),
                  "Whiskey Flats" = data %>% filter(ForestName == "WF"),
                  "Degrasse" = data %>% filter(ForestName == "DEG"))
    )
   
  
    if(study() == "SH")
      {data<- switch(input$species, 
                      "White-tailed Deer" = newData %>% filter(choice=="DEERWHITETAILED"| ForestName == "SH"),
                      "Chipmunk" = newData %>% filter(choice=="CHIPMUNK"| ForestName == "SH"),
                  "Coyote" = newData %>% filter(choice == "COYOTE"| ForestName == "SH"),
                  "Fisher" = newData %>% filter(choice == "FISHER"| ForestName == "SH"),
                  "Raccoon" = newData %>% filter(choice == "RACCOON"| ForestName == "SH"),
                  "Red Squirrel" = newData %>% filter(choice == "SQUIRRELRED"| ForestName == "SH"),
               "Gray Squirrel" = newData %>% filter(choice == "SQUIRRELGRAY"| ForestName == "SH"),
               "Black Bear" = newData %>% filter(choice == "BLACKBEAR"| ForestName == "SH"),
               "Red Fox" = newData %>% filter(choice == "FOXRED"| ForestName == "SH"),
               "Porcupine" = newData %>% filter(choice == "PORCUPINE"| ForestName == "SH"),
               "Bobcat" = newData %>% filter(choice == "BOBCAT"| ForestName == "SH"),
               "Weasel" = newData %>% filter(choice == "WEASEL"| ForestName == "SH"),
               "Striped Skunk" = newData %>% filter(choice == "SKUNKSTRIPED"| ForestName == "SH"),
               "Flying Squirrel" = newData %>% filter(choice == "SQUIRRELFLYING"| ForestName == "SH"),
               "Snowshoe Hare" = newData %>% filter(choice == "SNOWSHOEHARE"| ForestName == "SH"),
               "River Otter" = newData %>% filter(choice == "RIVEROTTER"| ForestName == "SH"),
               "Mink" = newData %>% filter(choice == "MINK"| ForestName == "SH"),
               "Other Small Mammal" = newData %>% filter(choice == "OTHERSMALLMAMMAL"| ForestName == "SH"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "SH"),)
      ggplot(data, aes("")) +
     geom_histogram(stat = "count", position = "dodge") +
     theme_bw() +
    xlab("Species") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    }
    
      if(study() == "BC" )
        {data<- switch(input$species, 
                      "White-tailed Deer" = newData %>% filter(choice=="DEERWHITETAILED"| ForestName == "BC"),
                      "Chipmunk" = newData %>% filter(choice=="CHIPMUNK"| ForestName == "BC"),
                  "Coyote" = newData %>% filter(choice == "COYOTE"| ForestName == "BC"),
                  "Fisher" = newData %>% filter(choice == "FISHER"| ForestName == "BC"),
                  "Raccoon" = newData %>% filter(choice == "RACCOON"| ForestName == "BC"),
                  "Red Squirrel" = newData %>% filter(choice == "SQUIRRELRED"| ForestName == "BC"),
               "Gray Squirrel" = newData %>% filter(choice == "SQUIRRELGRAY"| ForestName == "BC"),
               "Black Bear" = newData %>% filter(choice == "BLACKBEAR"| ForestName == "BC"),
               "Red Fox" = newData %>% filter(choice == "FOXRED"| ForestName == "BC"),
               "Porcupine" = newData %>% filter(choice == "PORCUPINE"| ForestName == "BC"),
               "Bobcat" = newData %>% filter(choice == "BOBCAT"| ForestName == "BC"),
               "Weasel" = newData %>% filter(choice == "WEASEL"| ForestName == "BC"),
               "Striped Skunk" = newData %>% filter(choice == "SKUNKSTRIPED"| ForestName == "BC"),
               "Flying Squirrel" = newData %>% filter(choice == "SQUIRRELFLYING"| ForestName == "BC"),
               "Snowshoe Hare" = newData %>% filter(choice == "SNOWSHOEHARE"| ForestName == "BC"),
               "River Otter" = newData %>% filter(choice == "RIVEROTTER"| ForestName == "BC"),
               "Mink" = newData %>% filter(choice == "MINK"| ForestName == "BC"),
               "Other Small Mammal" = newData %>% filter(choice == "OTHERSMALLMAMMAL"| ForestName == "BC"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "BC"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "BC"),)
      ggplot(data, aes("")) +
     geom_histogram(stat = "count", position = "dodge") +
     theme_bw() +
    xlab("Species") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
      }
         if(study() == "DON")
         { data<- switch(input$species, 
                      "White-tailed Deer" = newData %>% filter(choice=="DEERWHITETAILED"| ForestName == "DON"),
                      "Chipmunk" = newData %>% filter(choice=="CHIPMUNK"| ForestName == "DON"),
                  "Coyote" = newData %>% filter(choice == "COYOTE"| ForestName == "DON"),
                  "Fisher" = newData %>% filter(choice == "FISHER"| ForestName == "DON"),
                  "Raccoon" = newData %>% filter(choice == "RACCOON"| ForestName == "DON"),
                  "Red Squirrel" = newData %>% filter(choice == "SQUIRRELRED"| ForestName == "DON"),
               "Gray Squirrel" = newData %>% filter(choice == "SQUIRRELGRAY"| ForestName == "DON"),
               "Black Bear" = newData %>% filter(choice == "BLACKBEAR"| ForestName == "DON"),
               "Red Fox" = newData %>% filter(choice == "FOXRED"| ForestName == "DON"),
               "Porcupine" = newData %>% filter(choice == "PORCUPINE"| ForestName == "DON"),
               "Bobcat" = newData %>% filter(choice == "BOBCAT"| ForestName == "DON"),
               "Weasel" = newData %>% filter(choice == "WEASEL"| ForestName == "DON"),
               "Striped Skunk" = newData %>% filter(choice == "SKUNKSTRIPED"| ForestName == "DON"),
               "Flying Squirrel" = newData %>% filter(choice == "SQUIRRELFLYING"| ForestName == "DON"),
               "Snowshoe Hare" = newData %>% filter(choice == "SNOWSHOEHARE"| ForestName == "DON"),
               "River Otter" = newData %>% filter(choice == "RIVEROTTER"| ForestName == "DON"),
               "Mink" = newData %>% filter(choice == "MINK"| ForestName == "DON"),
               "Other Small Mammal" = newData %>% filter(choice == "OTHERSMALLMAMMAL"| ForestName == "DON"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "DON"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "DON"),)
      ggplot(data, aes("")) +
     geom_histogram(stat = "count", position = "dodge") +
     theme_bw() +
    xlab("Species") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
         }
           if(study() == "WF")
            {data<- switch(input$species, 
                      "White-tailed Deer" = newData %>% filter(choice=="DEERWHITETAILED"| ForestName == "WF"),
                      "Chipmunk" = newData %>% filter(choice=="CHIPMUNK"| ForestName == "WF"),
                  "Coyote" = newData %>% filter(choice == "COYOTE"| ForestName == "WF"),
                  "Fisher" = newData %>% filter(choice == "FISHER"| ForestName == "WF"),
                  "Raccoon" = newData %>% filter(choice == "RACCOON"| ForestName == "WF"),
                  "Red Squirrel" = newData %>% filter(choice == "SQUIRRELRED"| ForestName == "WF"),
               "Gray Squirrel" = newData %>% filter(choice == "SQUIRRELGRAY"| ForestName == "WF"),
               "Black Bear" = newData %>% filter(choice == "BLACKBEAR"| ForestName == "WF"),
               "Red Fox" = newData %>% filter(choice == "FOXRED"| ForestName == "WF"),
               "Porcupine" = newData %>% filter(choice == "PORCUPINE"| ForestName == "WF"),
               "Bobcat" = newData %>% filter(choice == "BOBCAT"| ForestName == "WF"),
               "Weasel" = newData %>% filter(choice == "WEASEL"| ForestName == "WF"),
               "Striped Skunk" = newData %>% filter(choice == "SKUNKSTRIPED"| ForestName == "WF"),
               "Flying Squirrel" = newData %>% filter(choice == "SQUIRRELFLYING"| ForestName == "WF"),
               "Snowshoe Hare" = newData %>% filter(choice == "SNOWSHOEHARE"| ForestName == "WF"),
               "River Otter" = newData %>% filter(choice == "RIVEROTTER"| ForestName == "WF"),
               "Mink" = newData %>% filter(choice == "MINK"| ForestName == "WF"),
               "Other Small Mammal" = newData %>% filter(choice == "OTHERSMALLMAMMAL"| ForestName == "WF"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "WF"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "WF"),)
      ggplot(data, aes("")) +
     geom_histogram(stat = "count", position = "dodge") +
     theme_bw() +
    xlab("Species") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
           }
            if(study() == "DEG")
              {data<- switch(input$species, 
                      "White-tailed Deer" = newData %>% filter(choice=="DEERWHITETAILED"| ForestName == "DEG"),
                      "Chipmunk" = newData %>% filter(choice=="CHIPMUNK"| ForestName == "DEG"),
                  "Coyote" = newData %>% filter(choice == "COYOTE"| ForestName == "DEG"),
                  "Fisher" = newData %>% filter(choice == "FISHER"| ForestName == "DEG"),
                  "Raccoon" = newData %>% filter(choice == "RACCOON"| ForestName == "DEG"),
                  "Red Squirrel" = newData %>% filter(choice == "SQUIRRELRED"| ForestName == "DEG"),
               "Gray Squirrel" = newData %>% filter(choice == "SQUIRRELGRAY"| ForestName == "DEG"),
               "Black Bear" = newData %>% filter(choice == "BLACKBEAR"| ForestName == "DEG"),
               "Red Fox" = newData %>% filter(choice == "FOXRED"| ForestName == "DEG"),
               "Porcupine" = newData %>% filter(choice == "PORCUPINE"| ForestName == "DEG"),
               "Bobcat" = newData %>% filter(choice == "BOBCAT"| ForestName == "DEG"),
               "Weasel" = newData %>% filter(choice == "WEASEL"| ForestName == "DEG"),
               "Striped Skunk" = newData %>% filter(choice == "SKUNKSTRIPED"| ForestName == "DEG"),
               "Flying Squirrel" = newData %>% filter(choice == "SQUIRRELFLYING"| ForestName == "DEG"),
               "Snowshoe Hare" = newData %>% filter(choice == "SNOWSHOEHARE"| ForestName == "DEG"),
               "River Otter" = newData %>% filter(choice == "RIVEROTTER"| ForestName == "DEG"),
               "Mink" = newData %>% filter(choice == "MINK"| ForestName == "DEG"),
               "Other Small Mammal" = newData %>% filter(choice == "OTHERSMALLMAMMAL"| ForestName == "DEG"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "DEG"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "DEG"),)
      ggplot(data, aes("")) +
     geom_histogram(stat = "count", position = "dodge") +
     theme_bw() +
    xlab("Species") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
            }
    
              if(study() == "WHIP")
             {data<- switch(input$species, 
                      "White-tailed Deer" = newData %>% filter(choice=="DEERWHITETAILED"| ForestName == "WHIP"),
                      "Chipmunk" = newData %>% filter(choice=="CHIPMUNK"| ForestName == "WHIP"),
                  "Coyote" = newData %>% filter(choice == "COYOTE"| ForestName == "WHIP"),
                  "Fisher" = newData %>% filter(choice == "FISHER"| ForestName == "WHIP"),
                  "Raccoon" = newData %>% filter(choice == "RACCOON"| ForestName == "WHIP"),
                  "Red Squirrel" = newData %>% filter(choice == "SQUIRRELRED"| ForestName == "WHIP"),
               "Gray Squirrel" = newData %>% filter(choice == "SQUIRRELGRAY"| ForestName == "WHIP"),
               "Black Bear" = newData %>% filter(choice == "BLACKBEAR"| ForestName == "WHIP"),
               "Red Fox" = newData %>% filter(choice == "FOXRED"| ForestName == "WHIP"),
               "Porcupine" = newData %>% filter(choice == "PORCUPINE"| ForestName == "WHIP"),
               "Bobcat" = newData %>% filter(choice == "BOBCAT"| ForestName == "WHIP"),
               "Weasel" = newData %>% filter(choice == "WEASEL"| ForestName == "WHIP"),
               "Striped Skunk" = newData %>% filter(choice == "SKUNKSTRIPED"| ForestName == "WHIP"),
               "Flying Squirrel" = newData %>% filter(choice == "SQUIRRELFLYING"| ForestName == "WHIP"),
               "Snowshoe Hare" = newData %>% filter(choice == "SNOWSHOEHARE"| ForestName == "WHIP"),
               "River Otter" = newData %>% filter(choice == "RIVEROTTER"| ForestName == "WHIP"),
               "Mink" = newData %>% filter(choice == "MINK"| ForestName == "WHIP"),
               "Other Small Mammal" = newData %>% filter(choice == "OTHERSMALLMAMMAL"| ForestName == "WHIP"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "WHIP"),
               "Opossum" = newData %>% filter(choice == "OPOSSUM"| ForestName == "WHIP"),)

      ggplot(data, aes(ForestName)) +
     geom_histogram(stat = "count", position = "dodge") +
     theme_bw() +
    xlab("Species") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
              }
    })

  }
shinyApp(ui = ui, server = server)
```

```{r}

    choices<-c("DEERWHITETAILED","MINK", "SKUNKSTRIPED","PORCUPINE")
    data<-newData %>% filter(choice %in% choices)
    study<-reactive(switch(input$forest,
             "South Hammond" = data %>% filter(ForestName=="SH"),
             "Beaver Creek" = data %>% filter(ForestName=="BC"),
             "Donnerville" = data %>% filter(ForestName == "DON"),
             "Whippoorwill Corners" = data %>% filter(ForestName == "WHIP"),
             "Whiskey Flats" = data %>% filter(ForestName == "WF"),
             "Degrasse" = data %>% filter(ForestName == "DEG"))
    )
    study<-data %>% filter(ForestName == "SH")
    ggplot(study, aes(choice)) +
          geom_histogram(stat = "count", position = "dodge") +
          theme_bw() +
          xlab("Species") +
          ylab("Number of Detections") +
          theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
  
```

