---
title: "Shiny App Forest Type"
output: html_notebook
---

This R Notebook will contain all the shiny apps broken down by forest type. Hoping to make tabs with visual results based on forest type. 


First, let's clear the environment and load the shiny package.
```{r}
rm(list = ls())
library(shiny)
library(dplyr)
library(readr)
library(ggplot2)
```

Now let's load the data frame we need. This will tell us the total number of species per forest and how many of each species were detected
```{r}
dataFinal <- read_csv("~/Documents/Kate_ThesisCode/dataFinal.csv")
```

Subset data to remove "Nothing Here" choice
```{r}
dump<-which(dataFinal$choice == "NOTHINGHERE")
newData<-dataFinal[-dump,]
```


Load the data that was modified to add trophic level column (for trophic output)
```{r}
mammals<- read_csv("mammals.csv")
mammals<- mammals[,2:16]
```

Load data and prepare data for covariates output
```{r}
Covariates <- read_csv("Covariates.csv")

Cov_sum<-Covariates %>% group_by(ForestType) %>% summarise(
  Water = mean(P_Water50), Mixed = mean(P_Mixed50), Evergreen = mean(P_Evergreen50), Deciduous = mean(P_Deciduous50), Development = mean(P_Development50), Agriculture = mean(P_Agriculture50), Barren = mean(P_Barren50), Shrub = mean(P_Shrub50), Wetland = mean(P_Wetland50), Herbaceous = mean(P_Herb50)
)

require(reshape2)
dat_molten <- melt(Cov_sum,"ForestType")
require(ggplot2)
```



Shiny app of species count by forest type
```{r}
ui <- fluidPage(

  # App title ----
  titlePanel("Species Count by Forest Type"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      selectInput("forest", h3("Choose your Forest Type"),
                  choices = c("Deciduous", "Pine"), selected = "Deciduous"),
        
       conditionalPanel( condition = "NOTHINGHERE",
                    checkboxInput("remove", "Remove Nothing Here"))

    ),

    # Main panel for displaying outputs ----
    mainPanel(  tabsetPanel(
        tabPanel("Species per Forest",
                 plotOutput(outputId = "foresthist")),
        tabPanel("Species Trophic Levels", plotOutput(outputId = "trophic")),
        tabPanel("Forest Composition", plotOutput(outputId = "covariates"))
      )

    )
  )
)

server <- function(input, output) {

  # Histogram of the number of species per forest type ----

  output$foresthist <- renderPlot({



if(input$remove){
  newdata<-switch(input$forest,
                "Deciduous" = newData %>% filter(ForestType=="Deciduous"),
                      "Pine" = newData %>% filter(ForestType=="Pine"))
  
    ggplot(newdata, aes(choice)) + 
     geom_histogram(stat = "count", position = "dodge") + 
     theme_bw() + 
    xlab("Species per Forest Type") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
}    
    else {
data<- switch(input$forest, 
                      "Deciduous" = dataFinal %>% filter(ForestType=="Deciduous"),
                      "Pine" = dataFinal %>% filter(ForestType=="Pine"))
          
      ggplot(data, aes(choice)) + 
     geom_histogram(stat = "count", position = "dodge") + 
     theme_bw() + 
    xlab("Species per Forest Type") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    }

})
   output$trophic <- renderPlot({

  data<-switch(input$forest, 
                      "Deciduous" = mammals %>% filter(ForestType=="Deciduous"),
                      "Pine" = mammals %>% filter(ForestType=="Pine"),)
  
    ggplot(data, aes(Trophic)) + 
     geom_histogram(stat = "count", position = "dodge") + 
     theme_bw() + 
    xlab("Trophic Level") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    
 
      })
   output$covariates <- renderPlot({

  data<-switch(input$forest, 
                      "Deciduous" = dat_molten %>% filter(ForestType=="Deciduous"),
                      "Pine" = dat_molten %>% filter(ForestType=="Pine"),
                  "Donnerville" = dat_molten %>% filter(Forest=="DON"),)
  
   ggplot(data, aes(x="", y=value, fill= variable)) + 
         geom_bar(stat="identity") + 
  coord_polar("y", start = 0) + 
     geom_text(aes(label = paste0(round(value), "%")), position = position_stack(vjust = 0.5))+
      theme(axis.line = element_blank(),
          plot.title = element_text(hjust=0.5)) + 
    labs(fill = "variable",
         x = NULL,
         y = NULL,
         title = "Forest Composition")
    
 
      })

    }


shinyApp(ui = ui, server = server)
```


Load the data that was modified to add trophic level column
```{r}
mammals<- read_csv("mammals.csv")
mammals<- mammals[,2:16]
```

Shiny app showing distribution of trophic hierarchy by forest type
```{r}

ui <- fluidPage(

  # App title ----
  titlePanel("Mammal Distribution by Forset Type"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      selectInput("forest", h3("Choose your Forest"),
                  choices = c("Deciduous", "Pine"), selected = "Deciduous"),
        
),
    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "trophic")

    )
  )
)

server <- function(input, output) {

  # Histogram of the number of different trophic levels per forest ----

  output$trophic <- renderPlot({

  data<-switch(input$forest, 
                      "Deciduous" = mammals %>% filter(ForestType=="Deciduous"),
                      "Pine" = mammals %>% filter(ForestType=="Pine"),)
  
    ggplot(data, aes(Trophic)) + 
     geom_histogram(stat = "count", position = "dodge") + 
     theme_bw() + 
    xlab("Trophic Level") +
     ylab("Number of Detections") +
     theme(axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5))
    
 
      })

    }



shinyApp(ui = ui, server = server)
```


Make a shiny app that breaks down the composition of each forest
```{r}
#First, load the data
Covariates <- read_csv("Covariates.csv")

Cov_sum<-Covariates %>% group_by(ForestType) %>% summarise(
  Water = mean(P_Water50), Mixed = mean(P_Mixed50), Evergreen = mean(P_Evergreen50), Deciduous = mean(P_Deciduous50), Development = mean(P_Development50), Agriculture = mean(P_Agriculture50), Barren = mean(P_Barren50), Shrub = mean(P_Shrub50), Wetland = mean(P_Wetland50), Herbaceous = mean(P_Herb50)
)

require(reshape2)
dat_molten <- melt(Cov_sum,"ForestType")
require(ggplot2)




ui <- fluidPage(

  # App title ----
  titlePanel("Forest Composition"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      selectInput("forest", h3("Choose your Forest"),
                  choices = c("Deciduous", "Pine"), selected = "Deciduous"),
        
),
    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "covariates")

    )
  )
)

server <- function(input, output) {

  # Pie chart of forest composition per forest ----

  output$covariates <- renderPlot({

  data<-switch(input$forest, 
                      "Deciduous" = dat_molten %>% filter(ForestType=="Deciduous"),
                      "Pine" = dat_molten %>% filter(ForestType=="Pine"),
                  "Donnerville" = dat_molten %>% filter(Forest=="DON"),)
  
   ggplot(data, aes(x="", y=value, fill= variable)) + 
         geom_bar(stat="identity") + 
  coord_polar("y", start = 0) + 
     geom_text(aes(label = paste0(round(value), "%")), position = position_stack(vjust = 0.5))+
      theme(axis.line = element_blank(),
          plot.title = element_text(hjust=0.5)) + 
    labs(fill = "variable",
         x = NULL,
         y = NULL,
         title = "Forest Composition")
    
 
      })

    }



shinyApp(ui = ui, server = server)
```
